Untitled

### 2019-11-12 Paper_resubmit.Rmd
regenerating figures for final paper

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(tidyverse)
library(BarrettsProgressionRisk)

@@

source('lib/load_patient_metadata.R')
#source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/BarrettsProgressionRisk/'
info.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info

patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')

@@

# Segment size: 5e06 with arms

## Quick look at our demographic information

```{r echo=F}

@@

We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.

```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme

p53Path = patient.info %>% dplyr::filter(Status == 'P' & p53.Status == 1) %>% dplyr::group_by(Pathology) %>% dplyr::summarise(n=length(Pathology)) %>% spread(Pathology,n)

p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)

@@

Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all. However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53. If we look at pathology the breakdown is even more obvious.  `r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.

`r p53.pred.byPath %>% kable(caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)') %>% kable_styling(full_width=F)`

# Data set up

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/models_5e6/50kb', sep='/') # QC'd

@@

load(file, verbose=F)
dim(dysplasia.df)
length(labels)

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}

chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

arm.loc = arrange(rbind(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = 
globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')# + theme(text=element_text(size=12))
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

@@

```{r loaddata, message=F, warning=F, echo=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']

x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

@@

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

@@

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

@@

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at `r select.alpha`. They share all non-zero features as well.
`r sapply(mp$stable.coeffients, length) %>% t %>% kable(caption='Total features at each value of alpha') %>% kable_styling(full_width = F)`

@@

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

@@

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}
perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')

```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

BarrettsProgressionRisk::showPredictionCalibration(noHGD.pred.confidence) + labs(subtitle='HGD excluded model')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

  fitNOHGD = models[[select.alpha]]

  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  colnames(pred) = 'Prediction'
  hgd.samples = base::merge(hgd.samples, pred, by.x='Samplename', by.y='row.names')

  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

  rm(no.hgd.plots, coefs, performance.at.1se, models)
}
#hist(hgd.samples$Prediction)

#rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  mutate_at(vars(start, end), list(as.numeric)) %>% mutate(chr = factor(chr,levels=c(1:22)))

features = features %>% mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), 
                    gl = factor(ifelse(coef > 0, 'gain','loss')))

features = left_join(features, chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p

ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

@@

@@

insilico.rocs[['discovery']] = roc

df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
)))

>>>>

---
title: "Untitled"
author: "Sarah Killcoyne"
date: "07/11/2019"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)

@@

source('lib/load_patient_metadata.R')
source('lib/common_plots.R')

dir = '~/Data/BarrettsProgressionRisk/'
info.dir = paste0(dir,'/QDNAseq')

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')

@@

# Demographics

@@

@@

# Copy number data set up

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/')

@@

load(file, verbose=F)
#dim(dysplasia.df)
#length(labels)

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

arm.loc = arrange(bind_rows(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))

globalCN.plot = cn.mtn.plot(globalCNdf %>% filter(seg.type != 'arms'), labeller(Status=n.status, chr=label_value))

cnLegend = get.legend(globalCN.plot)
globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')'))

globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

@@

## *Complexity Scores*

```{r loaddata, message=F, warning=F, echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']
complexity.score = tibble('Status'=labels, 'cx'=cx[names(labels)], 'Samplename' = names(labels)) %>%
  mutate(Status = factor(Status, labels = c('Non-Progressor','Progressor'))) %>%
  left_join(patient.info[,c('Samplename','Pathology','Patient')],by=c('Samplename'))

tt = t.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocX, title='Per sample cx') + plot.theme

@@

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose = F)
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

@@

mp = model.performance(performance.at.1se, coefs, folds, splits )
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

@@

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)
most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

@@

@@

@@

The initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, folds, splits)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot
ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

@@

@@

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

@@

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, folds, splits)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot
ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
```

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
```

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds', transpose=T)

@@

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=F)
```

@@

insilico.rocs[['discovery']] = roc
df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('AUC' = as.numeric(r$auc), 'Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
)))

#summary(filter(df, model == 'in-silico')$AUC)
